QPLIB: A Library of Quadratic Programming Instances

### 2018-02-03 AnswerToTheReferees.tex

\newcommand{\leo}[1]{{\color{red}#1}}
\newcommand{\fabio}[1]{{\color{cyan}#1}}
\newcommand{\emiliano}[1]{{\color{blue}#1}}   
\newcommand{\antonio}[1]{{\color{ForestGreen}#1}}     
\newcommand{\andrea}[1]{{\color{blue}{\bf #1}}}

\begin{document}

\section*{QPLIB: a Library of a Quadratic Programming Instances (MPC-022017-00003)-- Answer to the Referees}

\rep{
 Following the Area Editor suggestions, we 
 performed a revision of our manuscript. We carefully took into account all the comments made by the Referees and the Associate Editor.
 The new submitted document highlights in red the text that has been
added or changed.
In the following, the comments of the Associate Editor and of the Referees are reported, together with our point-by-point replies.
}

@@

{\it
I have received two reviews for this manuscript, and while there is consensus that the library makes an important contribution, one of the reviewers in particular had significant concerns with the library and the manuscript describing it. The most significant concern, which I share, is that the paper is significantly longer than it needs to be. In particular, I did not see a great need to review the solution methods as is done in Section 2.3. Mentioning the current solvers, and which classes of problems they can solve according to your taxonomy is perhaps useful for a reader's reference. But attempting to describe the algorithms themselves is somewhat futile, as it takes significant space and yet is still an inadequate description. For that, it would be better to simply provide some references for the benefit of readers who are interested in understanding the algorithms in the solvers. 
}

\rep{We agree with the AE that the description of the solution methods goes beyond the scope of the paper. For this reason we have streamlined Section 2.3 which now contains only a quick overview of the solvers and their solution methods. 
%\leo{Togliamo la spiegazione degli algoritmi e sostituiamola con una citazione di un book-tutorial-survey-introduction.}
}

{\it
I also agree with the reviewer's concern that some of the instance classes have only a single instance in them. If I am a researcher designing a method for solving instances in that class, I will be pretty disappointed to find that QPLIB only has a single relevant test instance. 
}

\rep{
We performed an internal call for instances among the members of the committee. 
We gathered $86$ new instances (belonging to the categories less that were not well represented) and we added them to the library. We added a macro classification of the instances, based on convexity and nature of the variables.
We believe that the majority of the algorithms used in the literature falls in one of these category.
This macro classification shows that the library is now well-balanced.
We kept also the detailed classification since it allows a fine grained analysis of the instances.
%%
%\leo{Sono d'accordo, ma come/dove troviamo le istanze supplementari? Sono le categorie con  $<10$ istanze nelle tavole 4 e 5, giusto? Dire che ce ne sono parecchie. 15 categorie su 23 hanno meno di 10 istanze (il numero 10 e` magico perche' le nostre mani hanno dieci dita, ovviamente).} 
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\begin{table}[h1]
%{
% \centering
% %\scriptsize
% \setlength{\tabcolsep}{18pt}
% \renewcommand \arraystretch{1.1}
%\begin{tabular}{llr}
%\toprule
%Variables & Convexity & \#\\
%\cmidrule(lr){1-3}
%%
% \textit{continuous}	& \textit{convex}		&  32 \\[1.2 ex]
% \textit{continuous}	& \textit{non-convex} 	&  102 \\[1.2 ex]
% \textit{discrete}	& \textit{convex}		&  31 \\[1.2 ex]
% \textit{discrete}	& \textit{non-convex}	&  288 \\[1.2 ex]
%\cmidrule(lr){1-3}
%Total		&				& 453\\
%%
%\bottomrule
%\end{tabular}
%\caption{Macro classification of the final set of instances}
%\label{tab:DDDDD}
%}
%\end{table}

@@

\rep{We would like to thank the Editor-in-Chief for having given us the possibility to perform a revision of our manuscript. In the following we present the improvements we made in accordance with the comments of the referees. }

\subsection*{Point-by-point reply to the Referee 1}

{\it
This paper presents a very thoughtfully constructed library of Quadratic Programming problems. I am not sure if there is a clear MPC policy on library
papers. I think it my merit a discussion at some time. Personally, I think libraries are very important so thoughtfully constructed ones such as this should
be encouraged and supported by MPC. However, I worry a bit about the over-
proliferation of such libraries, hence the reason for a possible discussion. Fortunately, I do not believe this is an issue with the library proposed in this
manuscript. The closest library to it I can think of is the CBLIB library I
mention bellow and the conceptual overlap is rather minimal. Furthermore, as
I believe the collection and selection of instances was very thorough and well
reasoned. Finally, I believe the brief survey and taxonomy of QP problems has
its own value beyond the library. I do have some comments bellow, but I believe
they can be easily addressed without the need for an additional detailed review.
Hence, my recommendation is the paper be accepted after a minor revision.
}

\rep{We thank the referee for his/her positive opinion on our manuscript. We hope that the new library will serve our scientific community.}

@@

\item I think it would be interesting to mention the Conic library CBLIB \url{http://cblib.zib.de} and its relation to the library described in the manuscript.

\rep{We updated the manuscript by adding the suggest link.}

\item Page 7, line -4. You say that convex means that the feasible region is
convex, but in page 19 you discuss the fact that SOCP problems that have
a convex feasible region could be considered nonconvex. Hence, maybe it
is more accurate to say that what you mean is that the constraint can be
converted to $f (x) \le c$ with $f$ convex by rearranging terms.

\rep{We removed a sentence from this paragraph to address this concern and we added a paragraph on this issue.
%TODO: Leo 
%\leo{secondo me il secondo commento si riferisce a pagina 19... a pagina 7 riga -4 non trovo niente su SOCP. Pero` sono d'accordo che a pagina 19 l'ultimo paragrafo della sezione 3.1 potrebbe essere migliorato. Cosi` e` un po' criptico. Perche' non inserire un bell'esempio di una forma bilineare nonconvessa (scrivendola esplicitamente come xQx e dando esplicitamente lo spettro di Q con la parte definita negativa) inserita in un vincolo che definisce una regione convessa? E poi io direi che facciamo la scelta di mettere queste istanze come "nonconvesse" perche' dimostrare convessita` in generale e` difficile (a questo proposito, io conosco i risultati del PhD di Amir Ali Ahmadi sulla difficolta` di riconoscere la convessita` in polinomi di grado $>= 4$, ma non ho mai visto una dimostrazione di difficolta` di riconoscere la convessita` della feasible region in vincoli quadratici; spero che qualcuno dei nostri co-autori ne sia al corrente)... Il referee ricorda che riconoscere i casi di forme quadratiche con una negative eigenvalue inserite in un vincolo convesso e` facile e  da` un metodo, di cui sarebbe bene ritrovare il primo caso in cui viene menzionato, oppure darne una breve dimostrazione... io non ne ero a conoscenza (e nemmeno mi viene in mente sui due piedi come dimostrarlo).} 
}

@@

\item Page 17. Doesn't CPLEX support some binary non-convex quadratic
problems.

\rep{We wrote an email to the development team of Cplex and they confirmed that their solver is able to tackle QGL and CGC 
% TODO:Fabio and Emiliano
%\leo{rispondiamo al referee che CPLEX e` classificato 'C' (complete) in tavola 2 rispetto a QGC e ricordiamogli che pero` CPLEX non risolve (a mia conoscenza e per il momento) QGQ. Tra l'altro, c'e` a p. 17 una menzione di "Table 2 below", tranne che la table 2 appare above. Diciamo solo "Table 2" senza il below e l'above, dato che la posizione di tavole e figure puo` comunque essere modificata sia da LaTeX sia dal typesetting finale.} 
}

@@

\end{enumerate}
}

\subsection*{Point-by-point reply to the Referee 2}

{\it
The paper presents a library of quadratic programming (QP) instances, where QP is broadly interpreted
to include quadratically constrained, and integer programs too. While I appreciate the overall goal of
the project, the paper is poorly written. It is often long-winded, and contains numerous factual errors,
and some omissions that must be corrected. Overall, the paper can easily be cut by 5-10 pages.
}

\rep{We followed the suggestion of the referee ad the AI and we have shortened unnecessary sections of the paper.}

@@

\item What do you mean by ``exact solution of the instances'' (p.10, l.-18)? I don't believe we can even
solve $Ax = b$ ``exactly'', except in some special cases. Please clarify this statement.

\rep{We added a comment about this issue in Section 2.1 
%\leo{possiamo inserire qualche nota sui modelli di computazione per i numeri reali, razionali e floating point: (a) real numbers con real RAM model, in pratica tutti i calcoli tra numeri reali sono O(1) (la gran parte del lavoro fatto da Blum, Shub e Smale e` sulla definizione di calcolabilita` e complessita` in questo 'ambiente'); 
%(b) hic sunt leones, non si sa nemmeno se la existential theory of rationals (cioe` risolvere sistemi polinomiali su $\mathbb{Q}$) sia decidibile o no [Pheidas 2000]; 
%(c) quello che la gente fa davvero: approssimare optimality e feasibility a $\epsilon>=0$ dato, nel migliore dei casi si dimostra anche che il tempo di calcolo o numero di iterazioni e` $O(1/poly(\epsilon)) o O(log(1/\epsilon))$.
%} 
}

\item The description of NLP solvers in Section 2.3.2 is at somewhat misleading. For example, KNITRO is also a primal-dual interior-point method, not just a barrier code (in fact primal-dual
methods are preferred). The description of the “active-set” methods is wrong, unless you refer
to the inner method, namely the active-set QP solves. Otherwise, SNOPT and KNITRO are
SQP methods, while CONOPT is a gradient-projection method, and MINOS is an augmented
Lagrangian method. You must correct this section

\rep{The complete description of the solver goes beyond the scope of the paper. For this reason we have streamlined the related Section.
%\leo{correggiamo gli errori e ringraziamo il referee} 
}

@@

\item Improve Fig.5: (1) What do the colors mean? (2) The plot on the right is not a sparsity pattern
of any Hessian, but at best the upper triangular sparsity pattern.

\rep{Corrected, thanks.}

\item I am concerned that problems are only available in GAMS, and tested on GAMS solvers. It would make the paper stronger if the authors also released AMPL model, for example, or made use of
JuMP.

\rep{
We thank the referee for having raised this delicate point. We definitely agree that formats are an issue.
For this reason we have translated all the instances also into the AMPL format and we have uploaded them on the website.
%\leo{Possiamo spiegare che proporre NOI le istanze in formati multipli darebbe la responsabilita` del traduttore a noi, che invece non vogliamo accettarla. D'altro canto penso che, sebbene sia stata una scelta discussa e effettuata democraticamente, non sia necessariamente la scelta giusta per ogni criterio. Per esempio, se il nostro criterio fosse la visibilita` della QPLIB (il numero di gente che la cita), dovremmo ASSOLUTAMENTE tradurre le istanze nel maggior numero possibile di formati. E se qualcuno ci fara` una correzione in futuro chi se ne frega, ringrazieremo e correggeremo. Francamente, se io fossi il dittatore della QPLIB, sceglierei di tradurre le istanze in quanti piu` formati sia possibile (almeno AMPL e SIF), e poi distribuirle con un warning.} \fabio{Adesso abbiamo anche aggiunto il formato AMPL, secondo me con questo il referee si mette tranquillo!} 
}

### 2018-02-03 QPLIB-3.tex
Update QPLIB-3.tex

We now analyse the features of the instances selected to be part of the
instance library. The characteristics of the instances are presented in Table
\ref{tab:DD} for \emph{discrete} instances
(\textit{*}\{\textit{B,M,I,G}\}\textit{*}) and in Table
\ref{tab:CC} for \emph{continuous} ones (\textit{*C*}).
For each category, the tables
report in column ``$\#$'' the corresponding number of instances. It can be seen
that the final set well respects the original distribution of the gathered
instances among the different categories. Indeed, the discrete categories
(\textit{LMQ}) or (\textit{QBL}) are well represented by $118$ and $59$
instances, respectively. Similarly, the continuous categories
(\textit{LCQ}) and (\textit{QCQ}) are well
represented by $50$ and $17$ instances, respectively. Moreover, the library
actually covers the large majority of all possible categories of instances.

@@

On the theme of nonconvexity, Figure~\ref{fig:pic_neg_eig} (left) focuses on
the instances with a quadratic objective function, ordered by the
the percentage ``hard'' eigenvalues in the Hessian $Q^0$ (vertical axis), by which
we mean negative eigenvalues in case of a minimization problem and positive eigenvalues in case of a maximization problem.
The instances are mostly clustered around two values. About 25\% of the instances have a convex (if minimization) or concave (if maximization) objective function,
i.e., they have 0\% of ``hard'' eigenvalues. Among the others, a vast
majority has around 50\% of ``hard'' eigenvalues. However, instances
with high or low percentages of ``hard'' eigenvalues are present, too.

\begin{figure}\centering
  \includegraphics[width=0.49\textwidth]{pic_neg_eig.pdf}
  \includegraphics[width=0.49\textwidth]{pic_density.pdf}
  \caption{``Hard'' eigenvalues (left) and density (right) of the Hessian $Q^0$ for QPLIB instances with a quadratic objective function. \label{fig:pic_neg_eig}}
\end{figure}

>>>>

\textcolor{red}{We now analyse the features of the instances selected to be part of the
 library. In Table \ref{tab:DDDDD}, we provide a global overview. The instances have been divided in \textit{continuous} vs \textit{discrete} and \textit{convex} vs \textit{non-convex}, forming in this way, a  classification of 4 macro categories. As previously mentioned, an instance is classified \textit{discrete} if it contains at least a binary or integer variable, and \textit{continuous} otherwise. On the other hand, an instance is classified as \textit{non-convex} if the objective function is non-convex and/or at least one of the constraints is  non-convex, and \textit{convex} otherwise. }

\begin{table}
\textcolor{red}{
 \centering
 %\scriptsize
 \setlength{\tabcolsep}{18pt}
 \renewcommand \arraystretch{1.1}
\begin{tabular}{llr}
\toprule
Variables & Convexity & \#\\
\cmidrule(lr){1-3}
%
 \textit{continuous}	& \textit{convex}		&  32 \\[1.2 ex]
 \textit{continuous}	& \textit{non-convex} 	&  102 \\[1.2 ex]
 \textit{discrete}	& \textit{convex}		&  31 \\[1.2 ex]
 \textit{discrete}	& \textit{non-convex}	&  288 \\[1.2 ex]
\cmidrule(lr){1-3}
Total		&				& 453\\
%
\bottomrule
\end{tabular}
\caption{Macro classification of the final set of instances}
\label{tab:DDDDD}
}
\end{table}

 The detailed characteristics of the instances are presented in Table
\ref{tab:DD} for \emph{discrete} instances
(\textit{*}\{\textit{B,M,I,G}\}\textit{*}) and in Table
\ref{tab:CC} for \emph{continuous} ones (\textit{*C*}).
For each category, the tables
report in column ``$\#$'' the corresponding number of instances. It can be seen
that the final set well respects the original distribution of the gathered
instances among the different categories. Indeed, the discrete categories
(\textit{LMQ}) or (\textit{QBL}) are well represented by $118$ and $59$
instances, respectively. Similarly, the continuous categories
(\textit{LCQ}) and (\textit{QCQ}) are well
represented by $50$ and $17$ instances, respectively. Moreover, the library
actually covers the large majority of all possible categories of instances.

@@

On the theme of nonconvexity, Figure~\ref{fig:pic_neg_eig} (left) focuses on
the instances with a quadratic objective function, ordered by 
 percentage of ``problematic'' \textcolor{red}{(defined using a tollerance of XXX)} \todo{please define the tolerance used to identify the ``Problematic'' eigenvalues and change the label of the plot}\ eigenvalues in the Hessian $Q^0$ (vertical axis), by which
we mean negative eigenvalues in case of a minimization problem and positive eigenvalues in case of a maximization problem.
The instances are mostly clustered around two values. About 25\% of the instances have a convex (if minimization) or concave (if maximization) objective function,
i.e., they have 0\% of ``problematic'' eigenvalues. Among the others, a vast
majority has around 50\% of ``problematic'' eigenvalues. However, instances
with high or low percentages of ``problematic'' eigenvalues are present, too.

\begin{figure}\centering
  \includegraphics[width=0.49\textwidth]{pic_neg_eig.pdf}
  \includegraphics[width=0.49\textwidth]{pic_density.pdf}
  \caption{``Problematic'' eigenvalues (left) and density (right) of the Hessian $Q^0$ for QPLIB instances with a quadratic objective function. \label{fig:pic_neg_eig}}
\end{figure}
